Preliminary Draft

### 2021-07-29 article/draft.tex
updates to YT section on recommendations

To examine this effect, we designed an experiment that simulates a user's behavior on YouTube. We used a list of $45$ videos marked as {\it False} by Science Feedback and classified under the category ``health'' (see section \ref{youtube_panels} for the classification). The experiment starts with one video marked as ``False'' from our list of videos and a clean watch history. For each video, our automated program collects around $20$ recommendations and clicks on the top $10$ videos in sequential order. After that, for every video in set of top $10$ recommendations, it collects their top $20$ recommended videos. In addition, to simulate a user's behavior, the automated program watches each video for a few minutes\footnote{If the video duration is ten minutes or less, the automated program watches the whole video. If the video duration is strictly higher than ten minutes, then the automated program watches only five minutes.}. Doing so, we collected $10$ $549$ YouTube recommended video entries, among which only $3$ $895$ were unique.

@@

To analyse the collected data, we ranked the recommended videos based on the number of times they got recommended. We then asked Science Feedback to fact-check the top 40 recommended videos, which amounts to $16\%$ of the data. Six videos (10.5\%) out of the top $40$ recommended videos were marked as misleading. This relatively small percentage points towards that idea that, YouTube is likely reducing the visibility of videos containing misinformation or alternatively maybe there is little misinformation on YouTube. Also it might be that the experiment setup did not influence the behavior of the recommendation engine, since we start with a single YouTube video marked as ``False'' every time. We further examined the top recommended channels in the collected data. We got $1$ $562$ different channels, and we counted the number of times every channel got recommended. Figure \ref{channels_yt} shows the top $10$ recommended channels, whose videos represent $30\%$ of the total set of recommended videos in our dataset. The top most recommended channel is Fox News, which has a ``mixed'' factual reporting level according to $iffy.news$ and has several failed fact-checks.\footnote{See \href{https://mediabiasfactcheck.com/fox-news-bias/}{https://mediabiasfactcheck.com/fox-news-bias/}.} Also in the top ten most recommended, there are two channels with a ``low'' factual reporting level: PrgerU and Sky news Australia.

\begin{figure}[h]
	\begin{center}
		\includegraphics[scale=0.5]{./img/channel_youtube.png} 
	\end{center}
	\caption{Top ten recommended channels.}
	\label{channels_yt}
\end{figure}

@@

In this article we have assessed three broad categories of interventions to tackle misinformation by Facebook, Twitter and YouTube: suspension of accounts, introduction of notices and flags and reducing the visibility of content. We have provided a methodology that can be useful for third-party monitoring. The code needed to collect the data and create the figures is available on the following public GitHub repository: \href{https://github.com/medialab/truth-and-trust-online-2021/}{https://github.com/medialab/truth-and-trust-online-2021/}, and we hope it can be a useful ressource for those working on addressing online misinformation.

@@

In the light of the present research, we discuss in this section data and policy related recommandations. Mainstream platforms such as Facebook, Twitter and Youtube are increasingly working towards higher data transparency (see links that redirect to the respective transparency centers in Appendix \ref{links}). Yet the present research sheds light on the lack and the splintering of data, which impede the academic community from studying successfully online misinformation and assessing the impact of platforms' interventions regarding this phenomenon. Hence, few comments are in order regarding the lack of data, the lack of transparency regarding misinformation related interventions, the splintering of data and the lack of interventions designed specifically for misinformation.

>>>>

To examine this effect, we designed an experiment that simulates a user's behavior on YouTube. We used a list of $45$ videos marked as {\it False} by Science Feedback and classified under the category ``health'' (see section \ref{youtube_panels} for the classification). The experiment starts with one video marked as ``False'' from our list of videos and a clean watch history. For each video, our automated program collects around\footnote{...} $20$ recommendations and clicks on the top $10$ videos in sequential order. After that, for every video in the set of the top $10$ recommendations, we collect their top $20$ recommended videos. In addition, to simulate a user's behavior, the automated program watches each video for a few minutes\footnote{If the video duration is ten minutes or less, the automated program watches the whole video. If the video duration is strictly higher than ten minutes, then the automated program watches only five minutes.}. Doing so, we collected $10$ $549$ YouTube recommended video entries, among which only $3$ $895$ were unique. 
To analyse the collected data, we look for misinformation: $(i)$ at the video level and $(ii)$ at the channel level.

@@

{\bf Video Level.} We ranked the recommended videos based on the number of times they got recommended. We then asked Science Feedback to fact-check the top 40 recommended videos.
Those $40$ videos were recommended $1$ $735$ times out of the $10$ $549$ YouTube recommended video entries (that is $16\%$ of the recommendations).
Among the top 40 recommended videos, four unique videos were marked as {\it misleading} and they were recommended $132$ times out of the $1$ $735$ recommendations (that is $7.7\%$ of the recommendations).
Four other videos were marked as {\it hyper-partisan} and they were recommended $184$ times ($10.7\%$).

{\bf Channel Level.} We examined the top recommended channels in the collected data. We got $1$ $562$ different channels, and we counted the number of times every channel got recommended. Figure \ref{channels_yt} shows the top $10$ recommended channels, whose videos represent $30\%$ of the total set of recommended videos in our dataset. Among the top three most recommended channels, there is Fox News and Sky News Australia, which both have a ``mixed'' factual reporting level according to {\it iffy.news} and have several failed fact-checks.\footnote{See \href{https://mediabiasfactcheck.com/fox-news-bias/}{https://mediabiasfactcheck.com/fox-news-bias/} and \href{https://mediabiasfactcheck.com/sky-news-australia/}{https://mediabiasfactcheck.com/sky-news-australia/}.} Also in the top ten most recommended, we find PragerU, which has a ``low'' level of factual reporting according to {\it iffy.news}.

At the video level, we start at each round with a YouTube video marked as ``False'' and we expected to have a high proportion of misinformation videos. But only {$18.4\%$} of the recommendations corresponded to $misleading$ or {\it hyper-partisan} videos.
This relatively small percentage hints towards that idea that YouTube is probably reducing the visibility of videos containing misinformation. However, when we investigated misinformation at the channel level, we found that videos belonging to channels that are not highly authoritative were being recommended. Finally, notice that the policy of reducing the visibility of misleading content is hard to prove on YouTube in general and even harder than on Facebook and Twitter.

\begin{figure}[h]
	\begin{center}
		\includegraphics[scale=0.6]{./img/channel_youtube.png} 
	\end{center}
	\caption{Top ten recommended channels among the $1$ $562$ different channels, that have posted the recommended videos in our dataset.}
	\label{channels_yt}
\end{figure}

@@

In this article we have assessed three broad categories of interventions to tackle misinformation by Facebook, Twitter and YouTube: suspension of accounts, introduction of notices and flags and reducing the visibility of content. We have provided a methodology that can be useful for third-party monitoring. The code needed to collect the data and create the figures is available on the following public GitHub repository: \href{https://github.com/medialab/truth-and-trust-online-2021/}{https://github.com/medialab/truth-and-trust-online-2021/}, and we hope it can be a useful ressource for researchers, NGOs and journalists addressing online misinformation.

@@

In the light of the present research, we discuss in this section data and policy related recommandations. Mainstream platforms such as Facebook, Twitter and Youtube are increasingly working towards higher data transparency (see links that redirect to the respective transparency centers in Appendix \ref{links}). Yet the present research sheds light on the lack and the splintering of data, which impede the academic community from studying successfully online misinformation and assessing the impact of platforms' interventions regarding this phenomenon. Hence, few comments are in order regarding the lack of data, the lack of transparency vis-\`{a}-vis misinformation related interventions, the splintering of data and the lack of interventions designed specifically for misinformation.
